---
title: "Evaluations of positive and negative stimuli using the Affective Misattribution Procedure (AMP) and self-reports"
subtitle: "Data processing"
author: "Template: Ian Hussey; content: [Student name]"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    code_folding: hide
    highlight: haddock
    theme: flatly
    toc: yes
    toc_float: yes
---

```{r, include=FALSE}

knitr::opts_chunk$set(message = FALSE,
                      warning = FALSE)

```

# Dependencies

```{r}

library(dplyr)
library(tidyr)
library(readr)
library(janitor) # for clean_names()
library(stringr)

```

# Get data

```{r}

# # demographics
# data_demographics_raw_messy <- read_csv("../data/raw/data_demographics_raw_messy.csv", skip = 2) |>
#   janitor::clean_names()

data_demographics_raw <- read_csv("../data/raw/data_demographics_raw.csv") |>
  janitor::clean_names()

# self report measure
data_selfreport_raw <- read_csv("../data/raw/data_selfreport_raw.csv") |>
  janitor::clean_names()

# affect attribution procedure
data_amp_raw <- read_csv("../data/raw/data_amp_raw.csv") |>
  janitor::clean_names()

```

# Demographics

Clean the demographics data so that summaries of age and gender can be made in the analysis.

```{r}

# wrangle the age data out of the raw
dat_age <- data_demographics_raw |>
  # rename for clarity
  rename(unique_id = subject, item = trialcode) |>
  # select useful columns
  select(unique_id, item, response) |>
  # retain rows that we need
  filter(item == "age") |>
  rename(age = response) |>
  # remove non-numbers or decimals from age
  mutate(age = str_remove_all(age, "[^\\d.]"),  # retains only digits and decimal points
         age = na_if(age, ""),
         age = as.numeric(age)) |>
  select(unique_id, age)
  #select(-item) # alternative: negative selection

# wrangle the gender data out of the raw
dat_gender <- data_demographics_raw |>
  # rename for clarity
  rename(unique_id = subject, item = trialcode) |>
  # select useful columns
  select(unique_id, item, response) |>
  # retain rows that we need
  filter(item == "gender") |>
  rename(gender = response) |>
  # remove non-numbers or decimals from age
  mutate(gender = str_to_lower(gender),
         gender = str_remove_all(gender, "[\\d.]"),
         gender = na_if(gender, ""), 
         gender = case_when(gender == "woman" ~ "female",
                            gender == "yes" ~ NA_character_,
                            gender == "man" ~ "male",
                            gender == "girl" ~ "female",
                            gender == "dude" ~ "male",
                            gender == "non binary" ~ "non-binary",
                            TRUE ~ gender)) |>
  select(unique_id, gender)

```

```{r}

# do we have duplicate data?
data_demographics_raw |>
  count(subject) |>
  arrange(desc(n))

# yes - n is different for some subjects

data_unique_id <- data_demographics_raw |>
  rename(unique_id = subject) |>
  count(unique_id) |>
  filter(n == 2) |>
  select(unique_id)
  
# multi-line typing: option-mouse drag on mac, alt-mouse drag on windows

set.seed(42)

data_unique_id_subset <- data_unique_id |>
  sample_n(size = 92)

write_csv(data_unique_id_subset, "data_unique_id_subset.csv")
  
```

## Combine

```{r}

# dat_gender |>
#   count(gender) |>
#   arrange(desc(n))
# 
# # dangerous - dont use as it assumes row order is equivalent
# # data_age_gender <- 
# #   bind_cols(dat_age,
# #             dat_gender)

data_age_gender <-
  full_join(x = dat_age,
            y = dat_gender, 
            by = "unique_id")

set.seed(42)
data_age_gender_subset <- data_age_gender |>
  sample_n(size = 90)

write_csv(data_age_gender_subset, "data_age_gender_subset.csv")

```

- in class: how to merge categories of "woman" and "female"?
- how to convert convert responses other than defined categories (e.g., "yes") to NA?

# Exclusions / data quality

## AMP

Create an exclusion variable `exclude_amp_performance` based on AMP performance: "exclude" if more than 10% of trials are < 100ms, or "include" if not.

Create an exclusion variable `exclude_amp_completeness` based on the AMP having the right number of trials.

```{r}

data_amp_performance_criteria <- data_amp_raw |>
  filter(blockcode == "test" & trialcode %in% c("prime_positive", "prime_negative")) |>
  select(unique_id = subject, prime = trialcode, correct, latency)

set.seed(42)

data_amp_summary_subset <- data_amp_performance_criteria |>
  group_by(unique_id) |>
  summarize(mean_reaction_time_behavioral_task = janitor::round_half_up(mean(latency), 0)) |>
  ungroup() |>
  sample_n(size = 31)

write_csv(data_amp_summary_subset, "data_amp_summary_subset.csv")
  
```

# Self-reports

Create a mean score for each participant of the three self-report items.

```{r}

set.seed(42)

data_selfreport_trial_level <- data_selfreport_raw |>
  select(unique_id = subject, item = trialcode, response) |> # note that you can do renames inside selects like this!
  filter(item %in% c("like", "positive", "prefer")) |>
  mutate(response = case_when(response == "Ctrl+'B'" ~ NA,
                              TRUE ~ response),
         response = as.numeric(response))

data_selfreport_summary_subset <- data_selfreport_trial_level |>
  group_by(unique_id) |>
  summarize(mean_self_report = janitor::round_half_up(mean(response), 2)) |>
  ungroup() |>
  sample_n(size = 76)

write_csv(data_selfreport_summary_subset, "data_selfreport_summary_subset.csv")

```

# Practicing joins

Using the data frames below and functions from the _join family, write code to do the following joins.  

create 'data_combined' by joining data_amp_summary_subset and data_age_gender_subset so that unique_ids in either data frame are retained. which join is this? implement it. 

create 'data_self_reports_and_their_amp_data' by joining data_selfreport_summary_subset and data_amp_summary_subset so that all participants have self-report data, + AMP data if available. which join is this? implement it. 

do the opposite: create 'data_amp_data_and_their_self_reports' by joining data_amp_summary_subset and data_selfreport_summary_subset so that all participants have AMP data, + self-report data if available. which join is this? implement it. 

join 'data_combined' and data_selfreport_summary_subset only unique_ids already present in data_combined are retained. which join is this? implement it. 

create 'data_missing_ids' which should list the unique_ids are missing from data_unique_id_subset but are present in at least one of data_age_gender_subset, data_amp_summary_subset, and data_selfreport_summary_subset. This will require two different joins. Which? Implement them.

```{r}

data_unique_id_subset <- read_csv("data_unique_id_subset.csv")
data_age_gender_subset <- read_csv("data_age_gender_subset.csv")
data_amp_summary_subset <- read_csv("data_amp_summary_subset.csv")
data_selfreport_summary_subset <- read_csv("data_selfreport_summary_subset.csv")


nrow(data_unique_id_subset)
nrow(data_age_gender_subset)
nrow(data_amp_summary_subset)
nrow(data_selfreport_summary_subset)

```




# Affect Misattribution Procedure

Create an overall AMP bias score. Score each trial as having been evalauted as congruent with the prime or not. eg When trialcode == "prime_positive" & evaluative_response == 1, then record as 1, etc. AMP score is the proportion of congruent responses.

```{r}

#data_amp_score_congruence <- data_amp_raw 

```

# Combine 

Combine data frames together to create one wide-format data frame where each row represents a participant.

Flag all duplicates for exclusion.

```{r}

# data_processed_before_exclusions 

```

# Define master exclusion variable

```{r}

#data_processed <- data_processed_before_exclusions

```

# Write to disk

```{r}

# # in case this dir doesn't exist, create it
# dir.create("../data/processed/")
# 
# # save data to disk in that dir
# write_csv(data_processed, "../data/processed/data_processed.csv")

```

# Session info

```{r}

sessionInfo()

```


